|

### 2022-10-11 paper/arxiv.Rmd
Version working with arxiv

For $d=2$ projections, the projection matrix is the sub-matrix $A$, of $O$ (an orthonormal basis for the p-dimensional space), formed by its first two columns as illustrated in Eq.~\ref{matrices}. Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}$, orthonormality of the projection matrix is expressed as $\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}$. Movement of the cursor takes the two components ${a_{m1},a_{m2}}$ into a selected new value ${a^*_{m1},a^*_{m2}}$. Although the motion is constrained by $a^{*2}_{m1}+a^{*2}_{m2}\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix. One possible algorithm to achieve this is
\begin{enumerate}
\item Cursor movement takes ${a_{m1},a_{m2}}\to {a^*_{m1},a^*_{m2}}$, called $V_m$ above.
\item The freedom to change the components $a_{m3}\cdots a_{mp}$ (the columns of $O$ not corresponding to the projection matrix $A$) is used to select a new orthonormal basis as follows:
\begin{enumerate}
\item For row $m$ one chooses $a^*_{m3}=\sqrt{1-a^{*2}_{m1}-a^{*2}_{m1}},~a^*_{m,k>3}=0$
\item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc, resulting in $A^{**}$.
\end{enumerate}
\item this results in the orthonormal basis $O^*$ and a new projection matrix with $A^{**}_{m1}=a^*_{m1},~A^{**}_{m2}=a^*_{m2}$.
\end{enumerate}
The random completion of $O$ outside the projection matrix provides an exploration of dimensions orthogonal to the projection plane. However, it renders projections in a way that is not continuous  and may be distracting. This can be alleviated by replacing the random completion with a rule restricting the size of the jumps, for example in step (b) the values $a_{i\neq m,j\geq3}$ can be left unchanged before orthogonalisation.

\begin{align}
O=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
\CodeBefore
       \columncolor{red!15}{1,2}
              \rowcolor{blue!15}{4}
     \Body
       & \stackrel{A_1}{\downarrow} & \stackrel{A_2}{\downarrow} &  & &\\
e_1\to & a_{11} & a_{12}  & a_{13} & \cdots & a_{1p} \\
e_2 \to& a_{21}  & a_{22}  & \cdot  &\cdots & a_{2p} \\
\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
 e_m \to& a_{m1}  & a_{m2}  & \cdot  &\cdots & a_{mp} \\   
\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
e_p \to&a_{p1} & a_{p2} &\cdot &\cdots & a_{pp} 
\end{pNiceArray},\quad (a_{m1},a_{m2})\to (a^*_{m1},a^*_{m2})
\label{matrices}
\end{align}

>>>>

For $d=2$ projections, the projection matrix is the sub-matrix $A$, of $O$ (an orthonormal basis for the p-dimensional space), formed by its first two columns as illustrated in Eq. (A1) . Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}$, orthonormality of the projection matrix is expressed as $\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}$. Movement of the cursor takes the two components ${a_{m1},a_{m2}}$ into a selected new value ${a^*_{m1},a^*_{m2}}$. Although the motion is constrained by $a^{*2}_{m1}+a^{*2}_{m2}\leq 1$, this is not sufficient to guarantee orthonormality of the new projection matrix. One possible algorithm to achieve this is
\begin{enumerate}
\item Cursor movement takes ${a_{m1},a_{m2}}\to {a^*_{m1},a^*_{m2}}$, called $V_m$ above.
\item The freedom to change the components $a_{m3}\cdots a_{mp}$ (the columns of $O$ not corresponding to the projection matrix $A$) is used to select a new orthonormal basis as follows:
\begin{enumerate}
\item For row $m$ one chooses $a^*_{m3}=\sqrt{1-a^{*2}_{m1}-a^{*2}_{m1}},~a^*_{m,k>3}=0$
\item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then proceeding as usual $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc, resulting in $A^{**}$.
\end{enumerate}
\item this results in the orthonormal basis $O^*$ and a new projection matrix with $A^{**}_{m1}=a^*_{m1},~A^{**}_{m2}=a^*_{m2}$.
\end{enumerate}
The random completion of $O$ outside the projection matrix provides an exploration of dimensions orthogonal to the projection plane. However, it renders projections in a way that is not continuous  and may be distracting. This can be alleviated by replacing the random completion with a rule restricting the size of the jumps, for example in step (b) the values $a_{i\neq m,j\geq3}$ can be left unchanged before orthogonalisation.

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.9\textwidth]{figures/matrix.png}}
\end{figure*}

<!-- 
\begin{align}
O=~~~& \begin{pNiceArray}{ccccc}[first-row,first-col]
\CodeBefore
       \columncolor{red!15}{1,2}
              \rowcolor{blue!15}{4}
     \Body
       & \stackrel{A_1}{\downarrow} & \stackrel{A_2}{\downarrow} &  & &\\
e_1\to & a_{11} & a_{12}  & a_{13} & \cdots & a_{1p} \\
e_2 \to& a_{21}  & a_{22}  & \cdot  &\cdots & a_{2p} \\
\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\ 
 e_m \to& a_{m1}  & a_{m2}  & \cdot  &\cdots & a_{mp} \\   
\vdots ~~~&\vdots & \vdots  & \vdots & \ddots & \vdots \\
e_p \to&a_{p1} & a_{p2} &\cdot &\cdots & a_{pp} 
\end{pNiceArray},\quad (a_{m1},a_{m2})\to (a^*_{m1},a^*_{m2})
\label{matrices}
\end{align}
-->
